---
title: 'Problem Set #3'
urlcolor: blue
output: 
  pdf_document: default
bibliography: ../bib/references.bib 
csl: ../bib/apa.csl
---

```{r setup, include = F}
knitr::opts_chunk$set(echo = F, warning = F, message = F, out.width = '75%', fig.align = 'center')


# Load libraries
library(tidyverse)
library(scales)
library(knitr)
library(kableExtra)


# Directory paths
getwd()
data_dir <- file.path(".", "files")
bib_dir <- file.path('..', 'bib')

input_data_dir <- file.path('..', 'data')
housing_df <- read_csv(file = file.path(input_data_dir, "housing.zip"))

```


# About the data
The data [@housing] contains information from the 1990 California census. @housing includes the following variables: longitude, latitude, housing_median_age, total_rooms, total_bedrooms, population, households, median_income, median_house_value, ocean_proximity . 

```{r}
list.files(data_dir)
housing_df <- readRDS(file.path(data_dir, "housing.RDS", fsep="/"))
housing_df %>%
      head(10) %>% 
      kable() %>%
      kable_styling(latex_options = c('scale_down', 'HOLD_position'))
```
# Data analysis

There is positive correlation between median income and median house value. Houses near bay or ocean have greater rate of change than inland houses.  
```{r}
housing_df %>% 
  ggplot(mapping = aes(x = median_income, y = median_house_value, color = ocean_proximity)) + 
  geom_point()+
  ggtitle('Correlation between Median Income and Median House Value') +
  xlab('Median income') +ylab('Median house value')+
  scale_color_brewer(palette = 'Spectral', name = "Ocean Proximity") +
  scale_x_continuous(labels = label_number(prefix = '$', suffix = 'K', scale = 10, accuracy = 1)) + 
  scale_y_continuous(labels = label_number(prefix = '$', suffix = 'K', scale = 1e-3, accuracy = 1)) +
  theme_minimal()
```

There is positive correlation between median income and median house value. Inland houses have the smallest  rate of change. There are limited island houses in this sample.  
```{r}
housing_df %>% 
  ggplot(mapping = aes(x = median_income, y = median_house_value)) + 
  geom_point()+
  facet_grid(rows = vars(ocean_proximity)) +
  ggtitle('Correlation between Median Income and Median House Value') +
  xlab('Median Income') +ylab('Median House Value') +
  scale_x_continuous(labels = label_number(prefix = '$', suffix = 'K', scale = 10, accuracy = 1)) + 
  scale_y_continuous(labels = label_number(prefix = '$', suffix = 'K', scale = 1e-3, accuracy = 1)) +
  theme_minimal()
```

This map shows a vague shape of the California landscape. There is a negative correlation between the logitude and latitude.
```{r}
housing_df %>% 
  ggplot(mapping = aes(x = longitude, y = latitude, color = ocean_proximity)) +
  geom_point()+
  ggtitle('Correlation between Longitude and Latitude') +
  xlab('Longitude') +ylab('Latitude') +
  scale_color_brewer(palette = 'Set3', name = "Ocean Proximity") +
  coord_fixed() +
  theme_light()
```

If the houses are in the same longitude, the lower the latitude, the higher the median house value. 
```{r}
housing_df %>% 
  ggplot(mapping = aes(x = longitude, y = latitude, color = median_house_value)) +
  geom_point()+
  ggtitle('Correlation between Longitude and Latitude by Median House Value') +
  xlab('Longitude') +ylab('Latitude') +
  scale_color_gradient(low = 'green', high = 'orange', name = "Median House Value", labels = label_number(prefix = '$', suffix = 'K', scale = 1e-3, accuracy = 1)) +
  coord_fixed() +
  theme_grey()
```

The average house value of island houses is more than 3 times higher than inland house, while houses near bay has slighltly higher average house values than houses near ocean. 
```{r}
housing_df %>% 
  group_by(ocean_proximity) %>% 
  summarize(average_house_value = mean(median_house_value, na.rm = TRUE)) %>% 
  ggplot(aes(x = ocean_proximity, y=average_house_value)) +
  geom_col(width = 0.6) +
  ggtitle('Average House Value by Ocean Proximity') +
  xlab('Ocean Proximity') +ylab('Average House Value') +
  scale_y_continuous(labels = label_number(prefix = '$', suffix = 'K', scale = 1e-3, accuracy = 1)) 
  
  
```
# References

